---
title: "Cousera - Reproducible Research - Course Project 1"
author: "András Hosznyák"
date: '2018-05-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading and preprocessing the data

The first step is to load the data from the activity.csv. The summary shows that there are many empty values on the steps column.

```{r}
activity <- read.csv("activity.csv")
summary(activity)
```

# What is mean total number of steps taken per day?

## The total number of steps taken per day

Before analyse the total number of steps taken per day, have to aggregate the data by days in a new variable. For the summary, the maximum value of x axis and the frequency breaks can be set.

```{r}
activity_day_total <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_day_total) <- c("date", "sum_of_step")
summary(activity_day_total)
```

## Histogram of the total number of steps taken each day

The histogram shows the frequency of the total number of steps taken each day.

```{r}
hist(activity_day_total$sum_of_step, main = "The total number of steps taken each day", breaks = 15, xlab = "Total steps by day", col = "red", ylim = c(0,20), xlim = c(0,25000) )
```

## The mean and median of the total number of steps taken per day

The median of the total number of steps taken per day is the following:

```{r}
median(activity_day_total$sum_of_step)
```

The mean of the total number of steps taken per day is the following:

```{r}
mean(activity_day_total$sum_of_step)
```

# What is the average daily activity pattern?

Before analyse the mean number of steps taken per interval, have to average the data by interval in a new variable. 

```{r}
activity_int_mean <- with(activity, aggregate(steps, by = list(interval), FUN = mean, na.rm = TRUE))
names(activity_int_mean) <- c("interval", "mean_of_step")
summary(activity_int_mean)
```

## Time series of the average number of steps

```{r}
plot(activity_int_mean$interval, activity_int_mean$mean_of_step, type = "l", lwd = 3, xlab = "Intervals (hhmm)", ylab = "Average step", main = "Time series of the average number of steps", col = "red")
```

## Which 5-minute interval contains the maximum number of steps?



```{r}
activity_int_mean[which.max(activity_int_mean$mean_of_step),]$interval
```

# Imputing missing values

There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

## The total number of rows with NA in the dataset

```{r}
na_index <- is.na(activity$steps)
sum(na_index)
```

## Devise a strategy for filling in all of the missing values in the dataset

The strategy does not need to be sophisticated, but I use the mean for that 5-minute interval. This value is 37.3826.

```{r}
na_fill_mean <- mean(activity_int_mean$mean_of_step)
na_fill_mean
```

## Create a new dataset

Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activity_fill <- activity
activity_fill[na_index, 1] <- na_fill_mean
head(activity_fill)
summary(activity_fill)
```

## Aggregate the data for the histogram

```{r}
activity_fill_day_total <- with(activity_fill, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_fill_day_total) <- c("date", "sum_of_step")
summary(activity_fill_day_total)
```

## Histogram of the total number of steps taken each day

The histogram shows the frequency of the total number of steps taken each day.

```{r}
hist(activity_fill_day_total$sum_of_step, main = "The total number of steps taken each day", breaks = 15, xlab = "Total steps by day", col = "red", ylim = c(0,20), xlim = c(0,25000) )
```

## The mean and median of the total number of steps taken per day

The median of the total number of steps taken per day is the following:

```{r}
median(activity_fill_day_total$sum_of_step)
```

The mean of the total number of steps taken per day is the following:

```{r}
mean(activity_fill_day_total$sum_of_step)
```

# Are there differences in activity patterns between weekdays and weekends?

## Create a new factor variable (weekday, weekend)

Creating a new factor variable - called weekday - With the weekdays and ifelse functions.

```{r}
activity_fill$day_name <- as.factor(weekdays(as.Date(activity_fill$date)))
activity_fill$weekday <- as.factor(ifelse(activity_fill$day_name == "szombat" | activity_fill$day_name == "vasárnap", "weekend", "weekday"))
```

## Time series plot for weekdays and weekends

The following plot shows the difference of the number of the average steps on weekdays or weekends.

```{r}
activity_fill_weekday_mean <- with(activity_fill, aggregate(steps ~ weekday + interval, data = activity_fill, FUN = mean))

library(ggplot2)
plot <- ggplot(activity_fill_weekday_mean, aes(x = interval, y = steps)) +
            geom_line(size = 1, color = "red") +
            facet_wrap( ~ weekday, nrow = 2) +
            labs(title = "Average steps by interval", x = "Interval", y = "Steps")

plot

```

